---
title: "p8105_hw1_yq2378"
author: "Qi Yumeng"
date: "2023-09-16"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

The dataset has 358 observations representing hourly measurements and 15 variables. According to help document, there are 5 time related variables like `year`, `month`, `day`, `hour` and `time_hour` (Date and hour of the recording as a POSIXct date). `origin` is the weather station. `temp`,`dewp` is temperature and dewpoint in F.`humid` is relative hunmidity. `wind_dir`, `wind_speed`, `wind_gust` is respectively wind direction (in degrees), speed and gust speed (in mph). `precip` is precipitation in millibars. `visib` is visbility in miles.

The mean temperature is around 39.58 F.

```{r load data}
x = c("moderndive", "ggplot2","ggpubr", "tidyverse")
lapply(x, require, character.only = TRUE)
data("early_january_weather")
```
```{r problem1}
# the size of the dataset (using nrow and ncol)
nrow(early_january_weather)
ncol(early_january_weather)
# the mean temperature
mean(early_january_weather$temp)
```

Over time, both temperature and humidity tend to increase. Of course, there are daily fluctuations, with temperatures typically peaking in the afternoon and dropping in the morning and at night.

```{r plots}
(p = ggplot(early_january_weather, 
            aes(x = time_hour, y =temp,color = humid))+geom_point() +
     scale_x_datetime(breaks = "3 day")+ 
                      theme_pubclean())
ggsave("output_plot1.png", plot = p, dpi = 300)
```

# Problem 2

```{r generate data frame}
# Set a seed for reproducibility
set.seed(2378)
# Create a random sample of size 10 from a standard Normal distribution
random_sample = rnorm(10, mean = 0, sd = 1)
# Create a logical vector indicating whether elements of the sample are greater than 0
logical_vector = random_sample > 0
# Create a character vector of length 10
character_vector = c("columbia", "168", "HYLO", "Mailman", "School", "COMOD", "Public", "Health", "AA", "Hammer")
# Create a factor vector of length 10 with 3 different factor "levels"
factor_vector = factor(sample(letters[1:3], 10, replace = TRUE))

# Create the data frame
df_problem2 = data.frame(
  RandomSample = random_sample,
  LogicalVector = logical_vector,
  CharacterVector = character_vector,
  FactorVector = factor_vector
)
# Print the data frame
print(data_frame)
```

When taking the mean of each variable, we get two warnings for `CharacterVector` and `FactorVector`. The other two variables, `RandomSample` and `LogicalVector` have means of their own.

```{r pull data}
df_problem2 %>% pull(RandomSample) %>% mean()
df_problem2 %>% pull(LogicalVector) %>% mean()
df_problem2 %>% pull(CharacterVector) %>% mean()
df_problem2 %>% pull(FactorVector) %>% mean()
```

The logical vector can be converted to numeric values following the mapping of True = 1 and False = 0

The character vector can be only converted to numeric values when the element is coated as character like "168". Otherwise, there will only be warnings and NAs.

The factor vector can be converted to numeric values based on the levels of the factor. The mapping is a = 1, b = 2 and c = 3 

```{r covert data, evals = FALSE}
as.numeric(df_problem2$LogicalVector)
as.numeric(df_problem2$CharacterVector)
as.numeric(df_problem2$FactorVector)
```
